\NAME: A framework for \\ Efficient Intrusion Detection\\ in the Internet of Things

### 2014-10-17 paper.tex
rewrote pattern analysis

In literature, many ID algorithm are described, but hardly any are implemented.
If implementations would be available, one would typically apply a pattern as
shown in listing \ref{alg:id-algo-application}.

@@

The implementations would be linked and called in a sequential way, when new
messages are observed on the network and at a regular interval to allow them to
do internal housekeeping. This includes processing the information extracted
from the network traffic, often in a statistical way.

This duality is a common design pattern of ID algorithms targeting MANETs. The
internals of the pattern reveal more common structures, as illustrated in
listing \ref{alg:id-algo-pattern}.

The parts of interest here are the loops at lines
\ref{alg:id-algo-pattern-loop1} and \ref{alg:id-algo-pattern-loop2}, the common
\emph{nodes} data on line \ref{alg:id-algo-pattern-common-data} and the
interaction with the network at lines \ref{alg:id-algo-pattern-send1} and
\ref{alg:id-algo-pattern-send2}.

@@

Given several ID algorithms, that all execute the same parsing loop, over the
same byte stream, performing comparable operations to find similar patterns,
one would clearly benefit from combining these loops, avoiding redundant
reevaluation of the same byte patterns.

The same goes for common data. Each algorithm not only has to parse the same
message, it also has to store information about nodes within its own scope,
resulting in a lot of duplicated information in memory.

The only way to solve this, would be to change the provided implementation,
pulling up the loops over the message and the nodes to a level above the loop
over the algorithms. But this comes at a development cost. This cost typically
isn't a one-time cost, e.g. in case a new version of the algorithm is released,
or simply when a different set of algorithms is selected due to a changing risk
analysis. Finally, changing, or simply implementing the algorithms from
scratch, also holds the risk of making mistakes.

@@

But there are hardly any implementations readily available. Probably the main
reason for this is the heterogeneity of the IoT. A prime example is the
\ttt{SEND} function. Different software stacks will have different application
programming interfaces (APIs) for accessing the network. Simple, macro-like
solutions can overcome this, but the dispersion of the calls to access the
network is causing the wireless radio to be on a lot more that one would hope
for.

A valid solution would try to combine messages in a single instance, reducing
the access to the network and even sharing some network overhead. The impact to
implement this manually will again outweigh the gain. An option would be to use
a software library, that provides a unified API for accessing the network and
collects individual calls and combines them in a single network message.

Besides the fact that all researchers would have to use the same library in a
consistent way, they would limit their operational scope to a single language,
serving only a fraction of the heterogeneous IoT.

\subsection{Situation Summary}

Starting from scratch allows optimizing the implementation of an IDS, but
doesn't scale. Reusing existing implementations also doesn't help, due to the
far from optimal sequential execution of the algorithms and the redundant
storage of common data. Software libraries can solve many of the issues, but
only cover parts of the entire spectrum, due to their tie-in with languages and
frameworks.

>>>>

We now extend the scope of the design pattern, to include its application. When
implementing different algorithms, one will typically encounter a pattern such
as describes in listing \ref{alg:id-algo-application}.

@@

The different implementations would be linked and called in a sequential way,
when new messages are observed on the network and at a regular interval to
allow them to do internal housekeeping.

Given these patterns, we can now identify the problem areas that tend to result
in suboptimal code when combining multiple ID algorithms.

@@

Two issues arise from the nested loops that originate from the combination of
the ID algorithm patterns and its application pattern. First, each algorithm,
by itself, will parse each observed message on the network, processing the same
byte stream, performing comparable operation to find similar patterns.

Second, there is common data: Each algorithm also has to store information
about nodes within its own scope, resulting in a lot of duplicated information
in memory.

Reimplementing the algorithms to eliminate these issues is not a valid option.
This development cost typically isn't a one-time cost, e.g. in case a new
version of the algorithm is released, or simply when a different set of
algorithms is selected due to a changing risk analysis. Finally, changing, or
simply implementing the algorithms from scratch, also holds the risk of making
mistakes.

@@

A third issue is hidden in the use of the network. The calls to the \ttt{SEND}
function are scattered around the algorithm, causing the wireless radio to be
accessed multiple times, keeping it active and energy consuming.

The heterogeneity of the IoT devices, results in many different software
stacks, each with their own network API. This also explains why there are very
few implementations readily available to developers, their is no common
platform to target.

Even using a unified software library would not only require a consistent
implementation of it, but would also limit the operational scope of the
research implementation, still restricting it to a single language.
